Verifying Cache Coherence Protocols for an Infinite Family of Cache Hierarchies

### 2014-02-06 paper/Introduction.tex
intro/system redundancy removed

Protocols, post verification, are often described via Term Rewrite Systems
(TRS) \adamc{citation for TRSes?}, which apply equational logic to describe
systems at an interesting point on the spectrum between declarative and
operational methods.  It was shown by Dave
\etal{}~\cite{DNA:CoherenceImplementation} that such a protocol, after its
verification, can be transliterated into Bluespec~\cite{Bluespec:TFRG}, a
hardware description language based on Guarded Atomic Actions, a restriction of
TRS~\cite{Hoe:TCAD}. Bluespec specifications can be directly implemented into
efficient hardware. In this way one reduces the correctness of the
implementation to the correctness of the protocol, with no need for separate
reasoning about an implementation. While the current Bluespec compiler does not
support some implementation elements that are common in practice (\eg{}
fairness guarantees, atomic actions running over multiple clock cycles), recent
work has provided key insights toward lifting the
restrictions~\cite{Karczmarek}. We believe that Bluespec and its future
evolutions represent a very promising platform for productive implementation of
hardware components, including cache-coherent shared memories, starting from a
description designed to support verification.

>>>>

Protocols, post verification, are often described via Term Rewrite Systems
(TRS) \adamc{citation for TRSes?}, which apply equational logic to describe
systems at an interesting point on the spectrum between declarative and
operational methods.  It was shown by Dave
\etal{}~\cite{DNA:CoherenceImplementation} that such a protocol, after its
verification, can be transliterated into Bluespec~\cite{Bluespec:TFRG}, a
hardware description language based on Guarded Atomic Actions, a restriction of
TRS~\cite{Hoe:TCAD}. Bluespec specifications can be directly implemented into
efficient hardware. In this way one reduces the correctness of the
implementation to the correctness of the protocol, with no need for separate
reasoning about an implementation. While the current Bluespec compiler does not
support some implementation elements that are common in practice (\eg{}
fairness guarantees, atomic actions running over multiple clock cycles), recent
work has provided key insights toward lifting the
restrictions~\cite{Karczmarek}. We believe that Bluespec and its future
evolutions represent a very promising platform for productive implementation of
hardware components, including cache-coherent shared memories, starting from a
specification designed to support verification.

### 2014-02-06 paper/System.tex
fixed added one line in system
intro/system redundancy removed

We first give an intuition on how the transitions work in the common case.
A cache can decide, spontaneously, to upgrade its state,
in which case it sends an upgrade request to its parent. The parent then makes
a local decision on whether to send a response to the requesting child or not
(based on its directory approximation, and its own state). If its own state is
lower the requested upgrade, then it cannot handle the request. The parent
must instead decide to upgrade its state -- the transition for doing so has
nothing to do with the incoming request.  Once the parent's state is not lower
than the requested upgrade, it makes sure that the rest of its children are
``compatible'' with the requested upgrade. If not, the parent must send
requests to the incompatible children to downgrade -- again the transition for
doing so has nothing to do with the incoming request. If the children are
compatible and its own state is higher, then the request can be handled and a
response is sent. Each of the transitions is explained in greater detail below.

@@

Note that the atomic transitions in Figure~\ref{trans} access only local states,
\ie{} a transition can read or write states only corresponding to a single cache
and/or the channel the cache is connected to. This restriction allows such a
system to be implemented directly into hardware. Bluespec System Verilog (BSV),
for instance, directly converts these transitions into efficient synchronous
hardware. Though these transitions logically happen one by one, BSV executes
several transitions simultaneously, as long as they do not access the same
state, \ie{} do not \emph{conflict}~\cite{Hoe:TCAD,HoeArvind:TRSSynthesis1}.
Sometimes, these transitions (both the guards and the actions) may not be
amenable to single-cycle implementations in hardware. For example, writing a
register array can actually take several hardware clock cycles, though
logically it is a single transition. Karczmarek \etal~\cite{Karczmarek} have
provided a scheme to convert atomic transitions into synchronous hardware in
which each transition can span several clock cycles.

>>>>

We first give an intuition on how the transitions work in the common case.
A cache can decide, spontaneously, to upgrade its state,
in which case it sends an upgrade request to its parent. The parent then makes
a local decision on whether to send a response to the requesting child or not
(based on its directory approximation, and its own state). If its own state is
lower the requested upgrade, then it cannot handle the request. The parent
must instead decide to upgrade its state -- the transition for doing so has
nothing to do with the incoming request.  Once the parent's state is not lower
than the requested upgrade, it makes sure that the rest of its children are
``compatible'' with the requested upgrade. If not, the parent must send
requests to the incompatible children to downgrade -- again the transition for
doing so has nothing to do with the incoming request. If the children are
compatible and its own state is higher, then the request can be handled and a
response is sent. The transitions for sending requests are allowed to happen
spontaneously without any trigger to model a family of protocols differing in
the policy for sending requests.

@@

Note that the atomic transitions in Figure~\ref{trans} access only local states,
\ie{} a transition can read or write states only corresponding to a single cache
and/or the channel the cache is connected to. This restriction allows such a
system to be implemented directly into hardware. Bluespec System Verilog (BSV),
for instance, directly converts these transitions into efficient synchronous
hardware. Though these transitions logically happen one by one, BSV executes
several transitions simultaneously, as long as they do not access the same
state, \ie{} do not \emph{conflict}~\cite{Hoe:TCAD,HoeArvind:TRSSynthesis1}.
%Sometimes, these transitions (both the guards and the actions) may not be
%amenable to single-cycle implementations in hardware. For example, writing a
%register array can actually take several hardware clock cycles, though
%logically it is a single transition. Karczmarek \etal~\cite{Karczmarek} have
%provided a scheme to convert atomic transitions into synchronous hardware in
%which each transition can span several clock cycles.
